On the Bit Complexity of Finding Points in Connected Components of a Smooth Real Hypersurface

### 2020-02-04 main.tex
typo
intro

% packages 
\usepackage[ruled,vlined]{algorithm2e}
\usepackage{mathrsfs}
\usepackage{bm}
\usepackage{enumitem}
% new commands and shortcuts
\def\td{{\bf todo}}
\def\bs{\textit{\textbf{s}}}
\def\bz{\textit{\textbf{0}}}
\def\Lb{\textit{\textbf{L}}}
\def\Xb{\textit{\textbf{X}}}
\def\Lambdab{\bm{\Lambda}}
\def\Thetab{\bm{\Theta}}
\def\thetab{\bm{\vartheta}}
\def\xb{\textit{\textbf{x}}}
\def\fb{\textit{\textbf{f}}}
\def\ab{\textit{\textbf{a}}}
\def\pb{\textit{\textbf{p}}}
\def\ajb{\overline{a_j}}
\def\bjb{\overline{b_{j,l}}}
\def\pjb{\overline{P_{j}}}
\DeclareMathOperator{\GL}{GL}
\DeclareMathOperator{\htt}{ht}
\def\C{\mathbb{C}}
\def\pr{\mathbb{P}}
\def\vt{\vartheta}
\def\Q{\mathbb{Q}}
\def\ZZ{\mathbb{Z}}
\def\R{\mathbb{R}}
\def\rs{\mathscr{R}}
\def\P{\mathscr{P}}
\def\d{\delta}
\def\sing{ \textrm{sing}}
\def\codim{ \textrm{codim}}
\def\jac{ \textrm{jac}}
\def\exp{\textrm{exp}}
\def\grad{\textrm{grad}}
\def\rank{\textrm{rank}}
\def\reg{\textrm{reg}}
\def\jt{\widetilde{J}}
\def\At{\widetilde{A}}
\def\Yt{\widetilde{Y}}
\def\kt{\widetilde{k}}
\def\dt{s}
%% \def\dt{\widetilde{d}}
\def\Dt{\widetilde{D}}
\def\rk{\textrm{rank }}
\def\pa{\partial}
\def\D{\Delta}
\def\Z{\frak{Z}}
\newcommand{\ZZ}{{\mathbb{Z}}}
\newcommand{\softO}{{O^{\sim}}}
\def\I{\frak{I}}
\def\Is{\frak{I}^{\star}}
\def\A{\frak{A}}
\def\fp{\frak{P}}
\def\sp{\mathscr{P}}
\def\la{\langle}
\def\ra{\rangle}
% matrices 
\def\scrQ{\ensuremath{\mathscr{Q}}}
\def\bbm{\begin{bmatrix}}
\def\ebm{\end{bmatrix}}
%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}

@@

\paragraph*{Background and problem statement.}
Computing one point in each connected component of a real algebraic
set $S$ is a basic subroutine in real alebraic and semi-algebraic
geometry; it is also useful in its own right, since it allows one to
decide if $S$ is empty or not. In this paper, we consider only the
case where $S$ is given as $S=V \cap \R^n$, where $V=V(f) \subset
\C^n$ is a complex hypersurface defined by a squarefree polynomial $f
\in \ZZ[X_1,\dots,X_n]$.

Algorithms for this task have been known for decades, and their
complexity is to some extent well understood. Suppose that $f$ has
degree $d$, and coefficients of bit-size $h$. Without making any
assumption on $f$, the algorithm given
in~\cite[Section~13.1]{BaPoRo03} solves our problem using $d^{O(n)}$
operations in $\Q$; in addition, the output of the algorithm is
represented by polynomials of degree $d^{O(n)}$, with coefficients of
bit-size $hd^{O(n)}$.  The key idea behind this algorithm goes back
to~\cite{GrVo88}: sample points are found through the computation of
critical points of well-chosen functions on $V(f)$.

@@

The starting point in our paper is the algorithm in~\cite{EMP}. In the
hypersurface case, its runtime is $d^{(4+o(1))n}$ operations in $\Q$;
as with many results in this line of work, the algorithm is
randomized: the algorithm needs to assume that we are in generic
coordinates; this is done by applying a random change of coordinates
prior to all computations. In addition, it relies on algorithms for
solving systems of polynomial equations that are themselves
randomized.  Altogether, this procedure chooses $n^{O(1)}$ random
vectors, each of them in an affine space of dimension $n^{O(1)}$;
every time a choice is made, failure corresponds to our chosen vector
lying in a hypersurface of the parameter space.

This paper revisits the algorithm in~\cite{EMP} (in the hypersurface
case) and gives a complete analysis of its probality of success and
its bit complexity.

\paragraph*{Data structures.} As in references such
as~\cite{GiHeMoPa95,GiHaHeMoMoPa97,GiHeMoMoPa98,BaGiHeMb97,EMP}, the
input polynomial $f$ is given by means of a {\em straight-line
  program}, that is, a sequence of operations $+,-,\times$ that
evaluates $f$, on the input variables $X_1,\dots,X_n$; the {\em
  length} of a straight-line program of such an objet is simply the
number of operations it performs. Any polynomial of degree $d$ in $n$
variables can be computed by a straight-line program of length $O\big
({{d+n} \choose n}\big)$ (evaluate all monomials of degree up to $d$
in $n$ variables, multiply them by their respective coefficients and
sum the result). However, some inputs may be given by shorter
straight-line program, and the algorithm is able to benefit from it.

The output of the algorithm is a finite set in $\C^n$ defined over
$\Q$. To represent it, we rely on a widely used data structure based
on univariate
polynomials~\cite{Kronecker82,Macaulay16,GiMo89,GiHeMoPa95,ABRW,GiHaHeMoMoPa97,GiHeMoMoPa98,Rouillier99}.
Consider a zero-dimensional algebraic set $S \subset \C^n$ defined
over $\Q$. A {\em zero-dimensional parametrization}
$\scrQ=((q,v_1,\dots,v_n),\lambda)$ of $S$ consists in polynomials
$(q,v_1,\dots,v_n)$, such that $q\in \Q[T]$ is monic and squarefree,
all $v_i$'s are in $\Q[T]$ and satisfy $\deg(v_i) < \deg(q)$, and in a
$\Q$-linear form $\lambda$ in variables $X_1,\dots,X_n$, such that
\begin{itemize}
\item $\lambda(v_1,\dots,v_N)=T q' \bmod q$, where $q'=\frac{\partial
  q}{\partial T}$;
\item we have the equality
  $S=\left \{\left(
      \frac{v_1(\tau)}{q'(\tau)},\dots,\frac{v_N(\tau)}{q'(\tau)}\right
    ) \ \mid \ q(\tau)=0 \right \}.$
\end{itemize}
The constraint on $\lambda$ says that the roots of $q$ are the values
taken by $\lambda$ on $S$. The parametrization of the coordinates by
rational functions having $q'$ as a denominator goes back
to~\cite{Kronecker82,Macaulay16}: as pointed out in~\cite{ABRW}, it
allows one to control precisely the size of the coefficients of
$v_1,\dots,v_n$.

\paragraph*{Main result.} To state our main result, we need to define 
the {\em height} of rationals, and polynomials with rational
coefficients.

The {\em height} of a non-zero rational $a=u/v$ is the maximum of
$\ln(|u|)$ and $\ln(|v|),$ where $u \in \mathbb{Z}$ and $v \in
\mathbb{N}$ are coprime. For a polynomial $f$ with rational
coefficients, if $v \in \mathbb N$ is the minimal common denominator
of all non-zero coefficients of $f$, then the \textit{height}
$\htt(f)$ of $f$ is defined as the maximum of the logarithms of $v$
and of the absolute values of the coefficients of $vf$.

\begin{theorem}
Suppose that $f\in\ZZ[X_1\hdots,X_n]$ is squarefree and satisfies $\deg
f \leq D, \htt(f) \leq s,$ and assume that $f$ is given by a
straight-line program $\Gamma$ of size $L$ with integer constants of
height at most $b.$ There exists a randomized algorithm that takes
$\Gamma, d,$ and $s$ as input and produces a zero-dimensional
parameterization of a set that contains at least one point in each
connected component of $V(f) \cap \R^n$ with probability at least
$1/2$. Otherwise the algorithm either produces a subset of the points
or FAIL. In any case, the algorithm uses
\[
O^{\sim}(kn(Lb+D^{2n}(s+D)(L+n^3D+n^4)))
\]
bit operations, with $k \geq -\lg(1-2^{(-1/n)})$. The polynomials in the output have degree at most $D^n,$ and height at most
\[
O^{\sim}((s+D)D^n + nD^n).
\]
\end{theorem}

Algorithms for deciding connectivity queries in real algebraic geometry
have been developed in a series of papers
\cite{EMP,SchostMohabBabySteps2011,SchostMohabBabySteps2014,TWT}. However,
these algorithms are based in the arithmetic complexity model, and bit
complexity questions have been left for future work. As bit complexity
provides a far more realistic model of computation compared to arithmetic
complexity,  which only counts each operation at unit cost and does not
account for coefficient size, our results therefore address an important
missing aspect of previous research.

Note that by the bit complexity we mean to count the number of all machine
operations require on a standard computer in terms of the length of a
standard representation (accounting for coefficient size, etc.). See below
for formal definitions.

The algorithm is Monte Carlo and can error on one side by producing a
proper subset of the points. Running the algorithm $k$ times gives a
list of outputs among which the highest cardinality set includes at
least one point in each connected component with probability at least
$\hdots.$

An \textit{algebraic set} $V \subset \C^n$ is the set of common zeros of a set of polynomials $\textit{\textbf{f}}=(f_1,\hdots,f_s)$ in $\C[X_1,\hdots,X_n].$

The \textit{dimension} of an algebraic set $V \subset \C^n$, denoted $\dim V,$ can be defined in the following ways:
\begin{enumerate}
    \item The number of generic hyperplanes needed to intersect with $V$ to obtain a finite set. 
    \item The Krull Dimension of $\C[X_1,\hdots,X_n]/I(V)$.
\end{enumerate}
The \textit{codimension} of $V$ is $n - \dim V$. An algebraic set is \textit{equidimensional} if each irreducible component has the same dimension.  If each component has dimension $d$ then we say it is $d-$equidimensional. 
The \textit{degree} of an algebraic set is the number of intersection points between itself and $\dim V$ generic hyperplanes.
\begin{example} 
An algebraic set of dimension zero is a finite set, with degree equal to its cardinality.
\end{example}
\begin{example}
An algebraic set of dimension 1 is a curve, with degree equal to the number of intersection points with a generic hyperplane.
\end{example}
\subsection{Critical / singular points}
\subsubsection{Regular and singular points of varieties}
Let $V \subset \C^n$ be a $d$-equidimensional algebraic set. 
The point $x \in V$ is a \textit{regular point} if $\dim (T_xV) = d.$ Otherwise $x$ is a \textit{singular point}. We let $\reg(V)$ and $\sing(V)$ respectively denote the regular and singular points of $V$. 
\par 
We assume that $V$ is smooth so that $V=\reg(V).$
\subsubsection{Critical points of polynomial mappings}
Let $\phi : V \rightarrow \C^m$ be a polynomial mapping. A \textit{critical point} of $\phi$ is a regular point of $V$ with $d_x \phi(T_xV) \not = \C^m,$ where $d_x \phi$ is the differential of $\phi$ at $x.$ A \textit{critical} value of $\phi$ is the image of a critical point by $\phi$.
\subsection{The zariski-tangent space}
Denote by $\grad_{\xb}(f)$ the evaluation of the gradient vector of $f$ in $\C[X_1,\hdots,X_n]$ at $\textit{\textbf{x}}$ in $\C$.
The \textit{Zariski-tangent space} to $V$ at $\textit{\textbf{x}} \in V$ is the vector space $T_{\xb}V$ defined by the equations $\grad_{\xb}(f) \cdot \textit{\textbf{v}}=0$ for all polynomials $f$ that vanish on $V$. For a polynomial system $\textit{\textbf{f}}=(f_1,\hdots,f_N)$ in $\C[X_1,\hdots,X_n]$, denote by $\jac(\textit{\textbf{f}})$ the Jacobian matrix.
The following is a direct consequence of \cite[Corollary 16.20]{ECA}.
If $V \subset \C^n$ is a $d$-equidimensional algebraic set with ideal $I(V)=\langle f_1,\hdots,f_N \rangle$, then at any point $\textbf{x}$ of $\reg(V),$ $\jac_\textbf{x}(\textbf{f})$ has full rank $n - \dim(V)$ and the kernel of $\jac_\textbf{x}(\textbf{f})$ is equal to $T_\textbf{x}V.$ 
\subsection{Polar varieties}
Let $V\subset \C^n$ be a smooth $d$-equidimensional variety. For $i \in \{1,\hdots,d\},$ denote by $\pi_i:\C^n \rightarrow \C^i$ the projection
$(x_1,\hdots,x_n) \mapsto  (x_1,\hdots,x_i)$. A \textit{critical point} $\textit{\textbf{x}}$ in $V$ is a singular point on $\pi_i.$ In other words, the point $\textit{\textbf{x}} \in V$ is a critical point if $\dim \pi_i( T_\textit{\textbf{x}}V) < i$. The $i$-th  \textit{polar variety} \[W(\pi_i,V) := \{\textit{\textbf{x}} \in V~|~\dim \pi_i(T_xV) < i\}\] is the set of critical points of $\pi_i$ on $V$. Denote by $\I$ the ideal \[\frak{I}(\pi_i,V) := I(W(\pi_i,V)).\] 
Consider $f \in \mathbb{Q}[X_1,\hdots,X_n]$ squarefree with $V(f)$ smooth. Then, for $i \in \{1,\hdots,n\}$, the polar variety $W(\pi_i,V)$ is defined by the vanishing of \[f,\frac{\pa f}{\pa X_{i+1}},\hdots,\frac{\pa f}{\pa X_n}.\]
We have a proof but maybe we should reference it?
\subsection{Changes of variables}
Consider $A \in \C^{n^2}, f \in \C[X_1,\hdots,X_n], \textit{\textbf{x}} \in \mathbb{C}^n,$ and $V \subset \C^n$. We denote by 
$V^A$ the set $\{x^A~|~\xb \in V\}, f^A$ the polynomial $f(A \xb)$ and $\xb^A$ denotes $A^{-1}\xb$. Notice $f^A(\xb^A) = f(A A^{-1}\xb)=f(\xb).$
\section{Genericity properties}
Consider $f \in \mathbb{Q}[X_1,\hdots,X_n]$ with total degree $D$.  
We say that $f$ satisfies $\textbf{H}$ if 
\begin{enumerate}
    \item $f$ is squarefree;
    \item $V(f)$ is smooth.
\end{enumerate}
An equidimensional algebraic set $X$ of dimension $d$ is in \textit{Noether position} for $\pi_d$ when the extension 
\C[X_1,\hdots,X_d] \rightarrow \C[X_1,\hdots,X_n]/I(V)
is injective and integral.  Then, for any $\xb \in \C^d,$ the fiber 
$V \cap \pi_d^{-1}(\xb)$ has dimension zero. For $i\in\{1,\hdots,n\}$, and for a randomly chosen point $p_{i-1} = (x_1,\hdots,x_{i-1}) \in \Q^{i-1}$, we say that $f$ satisfies $\textbf{H}_i^{'}$ if 
\begin{enumerate}
\item $\I(\pi_i,V)$ is radical;
\item either $W(\pi_i,V)$ is empty or $(i-1)$-equidimensional;
\item $W(\pi_i,V)$ is smooth;
\item $\langle X_1-x_1,\hdots,X_{i-1}-x_{i-1}, f,  \frac{\pa f}{\pa X_{i+1}},\hdots,\frac{\pa f}{\pa X_n}\rangle $ is radical; 
\item $V(X_1-x_1,\hdots,X_{i-1}-x_{i-1}, f,  \frac{\pa f}{\pa X_{i+1}},\hdots,\frac{\pa f}{\pa X_n})$ is finite;
\item $V(X_1-x_1,\hdots,X_{i-1}-x_{i-1}, f,  \frac{\pa f}{\pa X_{i+1}},\hdots,\frac{\pa f}{\pa X_n})$ is smooth;
\item either $W(\pi_i,V)$ is empty or in Noether position for $\pi_{i-1}$.
\end{enumerate}
Suppose that $f \in \Q[X_1,\hdots,X_n]$ satisfies \textbf{H}. Then there exists a Zariski closed set $\D \subset \mathbb{C}^{n^2}$ of degree at most $D^{\textrm{\td}}$ such that, if $\textbf{A}$ is chosen from  $\C^{n^2}-\Delta$ ( $\Q^{n^2}-\Delta$?) then $f^A$ satisfies $\textbf{H}_i^{'}$ for all $i \in \{1,\hdots,n\}.$
\begin{corollary} 
Fix $S \subset \mathbb{Q}$ with $|S| \geq \epsilon^{-1} D^{\textrm{\td}}$ and $\epsilon > 0$. Then for $A\in S^{n^2}$ chosen randomly, the probability that $f^A$ satisfies $H_i^{'}$ for all $i \in \{1,\hdots,n\}$ is at least $1-\epsilon.$
\end{corollary}
\section{Weak Transversality}

>>>>

% packages 
\usepackage[ruled,vlined]{algorithm2e}
\usepackage{mathrsfs}
\usepackage{bm}
\usepackage{enumitem}
% new commands and shortcuts
\def\td{{\bf todo}}
\def\bs{\textit{\textbf{s}}}
\def\bz{\textit{\textbf{0}}}
\def\Lb{\textit{\textbf{L}}}
\def\Xb{\textit{\textbf{X}}}
\def\Lambdab{\bm{\Lambda}}
\def\Thetab{\bm{\Theta}}
\def\thetab{\bm{\vartheta}}
\def\xb{\textit{\textbf{x}}}
\def\fb{\textit{\textbf{f}}}
\def\ab{\textit{\textbf{a}}}
\def\pb{\textit{\textbf{p}}}
\def\ajb{\overline{a_j}}
\def\bjb{\overline{b_{j,l}}}
\def\pjb{\overline{P_{j}}}
\DeclareMathOperator{\GL}{GL}
\DeclareMathOperator{\htt}{ht}
\def\C{\mathbb{C}}
\def\pr{\mathbb{P}}
\def\vt{\vartheta}
\def\Q{\mathbb{Q}}
\def\R{\mathbb{R}}
\def\rs{\mathscr{R}}
\def\P{\mathscr{P}}
\def\d{\delta}
\def\sing{ \textrm{sing}}
\def\codim{ \textrm{codim}}
\def\jac{ \textrm{jac}}
\def\exp{\textrm{exp}}
\def\grad{\textrm{grad}}
\def\rank{\textrm{rank}}
\def\reg{\textrm{reg}}
\def\jt{\widetilde{J}}
\def\At{\widetilde{A}}
\def\Yt{\widetilde{Y}}
\def\kt{\widetilde{k}}
\def\dt{s}
%% \def\dt{\widetilde{d}}
\def\Dt{\widetilde{D}}
\def\rk{\textrm{rank }}
\def\pa{\partial}
\def\D{\Delta}
\def\Z{\frak{Z}}
\newcommand{\ZZ}{{\mathbb{Z}}}
\newcommand{\softO}{{O^{\sim}}}
\def\I{\frak{I}}
\def\Is{\frak{I}^{\star}}
\def\A{\frak{A}}
\def\fp{\frak{P}}
\def\sp{\mathscr{P}}
\def\la{\langle}
\def\ra{\rangle}
% matrices 
\def\scrQ{\ensuremath{\mathscr{Q}}}
\def\bbm{\begin{bmatrix}}
\def\ebm{\end{bmatrix}}
%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}

@@

\paragraph*{Background and problem statement.}
Computing one point in each connected component of a real algebraic
set $S$ is a basic subroutine in real alebraic and semi-algebraic
geometry; it is also useful in its own right, since it allows one to
decide if $S$ is empty or not.

In this paper, we consider the case where $S$ is given as $S=V \cap
\R^n$, where $V=V(f) \subset \C^n$ is a complex hypersurface defined
by a squarefree polynomial $f \in \ZZ[X_1,\dots,X_n]$.
Algorithms for this task have been known for decades, and their
complexity is to some extent well understood. Suppose that $f$ has
degree $d$, and coefficients of bit-size $h$. Without making any
assumption on $f$, the algorithm given
in~\cite[Section~13.1]{BaPoRo03} solves our problem using $d^{O(n)}$
operations in $\Q$; in addition, the output of the algorithm is
represented by polynomials of degree $d^{O(n)}$, with coefficients of
bit-size $hd^{O(n)}$.  The key idea behind this algorithm goes back
to~\cite{GrVo88}: sample points are found through the computation of
critical points of well-chosen functions on $V(f)$.

@@

Our starting point is the algorithm in~\cite{EMP}. In the hypersurface
case, its runtime is $d^{(4+o(1))n}$ operations in $\Q$.  As with many
results in this vein, the algorithm is randomized: we need to assume
that we are in generic coordinates; this is done by applying a random
change of coordinates prior to all computations. In addition, the
algorithm relies on procedures for solving systems of polynomial
equations that are themselves randomized.  Altogether, we choose
$n^{O(1)}$ random vectors, each of them in an affine space of
dimension $n^{O(1)}$; every time a choice is made, there exists a
hypersurface of the parameter space that one has to avoid in order to
guarantee success. In this paper, we revisit this algorithm and give a
complete analysis of its probality of success and its bit complexity.

\paragraph*{Data structures.} As in references such
as~\cite{GiHeMoPa95,GiHaHeMoMoPa97,GiHeMoMoPa98,BaGiHeMb97,EMP}, the
input polynomial $f$ is given by means of a {\em straight-line
  program}, that is, a sequence of operations $+,-,\times$ that takes
as input $X_1,\dots,X_n$ and evaluates $f$; the {\em length} of a
straight-line program is simply the number of operations it performs.

Any polynomial of degree $d$ in $n$ variables can be computed by a
straight-line program of length $O\big ({{d+n} \choose n}\big)$:
evaluate all monomials of degree up to $d$ in $n$ variables, multiply
them by their respective coefficients and sum the result. However,
some inputs may be given by shorter straight-line program, and the
algorithm is able to benefit from it.

The output of the algorithm is a finite set in $\overline{\Q}{}^n$. To
represent it, we rely on a widely used data structure based on
univariate
polynomials~\cite{Kronecker82,Macaulay16,GiMo89,GiHeMoPa95,ABRW,GiHaHeMoMoPa97,GiHeMoMoPa98,Rouillier99}.
Consider a zero-dimensional algebraic set $S \subset \C^n$ defined
over $\Q$. A {\em zero-dimensional parametrization}
$\scrQ=((q,v_1,\dots,v_n),\lambda)$ of $S$ consists in polynomials
$(q,v_1,\dots,v_n)$, such that $q\in \Q[T]$ is monic and squarefree,
all $v_i$'s are in $\Q[T]$ and satisfy $\deg(v_i) < \deg(q)$, and in a
$\Q$-linear form $\lambda$ in variables $X_1,\dots,X_n$, such that
\begin{itemize}
\item $\lambda(v_1,\dots,v_n)=T q' \bmod q$;
\item we have the equality
  $S=\left \{\left(
      \frac{v_1(\tau)}{q'(\tau)},\dots,\frac{v_n(\tau)}{q'(\tau)}\right
    ) \ \mid \ q(\tau)=0 \right \}.$
\end{itemize}
The constraint on $\lambda$ says that the roots of $q$ are the values
taken by $\lambda$ on $S$. The parametrization of the coordinates by
rational functions having $q'$ as a denominator goes back
to~\cite{Kronecker82,Macaulay16}: as pointed out in~\cite{ABRW}, it
allows one to control precisely the size of the coefficients of
$v_1,\dots,v_n$.

\paragraph*{Main result.} To state our main result, we need to define 
the {\em height} of a rational number, and of a polynomial with
rational coefficients.

The {\em height} of a non-zero $a=u/v \in \Q$ is the maximum of
$\ln(|u|)$ and $\ln(v),$ where $u \in \mathbb{Z}$ and $v \in
\mathbb{N}$ are coprime. For a polynomial $f$ with rational
coefficients, if $v \in \mathbb N$ is the minimal common denominator
of all non-zero coefficients of $f$, then the \textit{height}
$\htt(f)$ of $f$ is defined as the maximum of the logarithms of $v$
and of the absolute values of the coefficients of $vf$.

\begin{theorem}
  Suppose that $f\in\ZZ[X_1\hdots,X_n]$ is squarefree, satisfies
  $\deg(f) \leq d, \htt(f) \leq s$, and that $V(f) \subset \C^n$ is
  smooth.  Assume also that $f$ is given by a straight-line program
  $\Gamma$ of size $L$ with integer constants of height at most $b.$

  There exists a randomized algorithm that takes $\Gamma, d,$ and $s$
  as input and produces a zero-dimensional parameterization of a set
  that contains at least one point in each connected component of
  $V(f) \cap \R^n$ with probability at least $1/2$. Otherwise the
  algorithm either produces a subset of the points or FAIL. In any
  case, the algorithm uses
  \[
  O^{\sim}(k(Lb+d^{2n}(s+d)(L+d)))
  \]
  bit operations, with $k \geq -\lg(1-2^{(-1/n)})$. The polynomials in
  the output have degree at most $d^n,$ and height $
  O^{\sim}((s+d)d^n).$
\end{theorem}
The algorithm itself is rather simple. To describe it, we need to
define {\em polar varieties}, which will play a crucial role in this
paper. Let $V=V(f)$, for $f$ as in the theorem. For $i \in
\{1,\hdots,n-1\},$ denote by $\pi_i:\C^n \rightarrow \C^i$ the
projection $(x_1,\hdots,x_n) \mapsto (x_1,\hdots,x_i)$.  The $i$-th
\textit{polar variety} \[W(\pi_i,V) := \{\xb \in V~|~\dim \pi_i(T_\xb
V) < i\}\] is the set of critical points of $\pi_i$ on $V$.  It is
thus defined by the vanishing of \[f,\frac{\pa f}{\pa
  X_{i+1}},\hdots,\frac{\pa f}{\pa X_n}.\]

In general, we cannot say much about the geometric properties of
$W(\pi_i,V)$, but if we apply a generic change of coordinates $\bm A$,
$W(\pi_i,V)$ is then known to be equidimensional of dimension $(i-1)$
or empty~\cite{BaGiHeMb97}, and to be in so-called {\em Noether
  position}~\cite{EMP} (background notions in algebraic geometry are
in~\cite{Mumford76,Shafarevich77,Eisenbud96}; we will recall key
definitions). If this is case, it suffices to choose arbitrary
$\alpha_1,\dots,\alpha_{n-1}$ in $\Q$, and solve the systems defined
by
\begin{equation}\label{eq:syst}
X_1-\alpha_1,\dots,X_{i-1}-\alpha_{i-1}, f, \frac{\pa f}{\pa
  X_{i+1}} ,\dots,\frac{\pa f}{\pa X_n},  
\end{equation}
for $i=1,\dots,n$.  They all admit finitely many solutions, and
Theorem~2 in~\cite{EMP} proves that the union of their solution sets
contains one point on each connected component of $V \cap \R^n$.

Our main contribution is to analyse precisely what conditions on our
change of coordinates $\bm A$ guarantee success. This is done by
revisiting the key ingredients in the proofs given
in~\cite{BaGiHeMb97} and~\cite{EMP} and giving quantitative versions
of these results, bounding the degree of the hypersurfaces we have to
avoid.  To solve the equations~\eqref{eq:syst}, we use the algorithm
in~\cite{SH}, for which a complete bit complexity analysis is
available.

This work should be seen as a first step toward the analysis of
randomized algorithms in real algebraic geometry. An immediate
follow-up question would be to handle the case of algebraic sets
defined by {\em regular sequences}: the algorithm in~\cite{EMP} still
applies, but the modifications needed would take us to far for the
current format. Further yet, randomized algorithms for deciding {\em
  connectivity queries} on smooth, compact algebraic sets have been
developed in a series of papers
\cite{SchostMohabBabySteps2011,SchostMohabBabySteps2014}, and could be
revisited using the techniques introduced here.

%% The \textit{dimension} of an algebraic set $V \subset \C^n$, denoted $\dim V,$ can be defined in the following ways:
%% %
%% \begin{enumerate}
%%     \item The number of generic hyperplanes needed to intersect with $V$ to obtain a finite set. 
%%     \item The Krull Dimension of $\C[X_1,\hdots,X_n]/I(V)$.
%% \end{enumerate}
%% %
%% The \textit{codimension} of $V$ is $n - \dim V$. An algebraic set is \textit{equidimensional} if each irreducible component has the same dimension.  If each component has dimension $d$ then we say it is $d-$equidimensional. 
%% The \textit{degree} of an algebraic set is the number of intersection points between itself and $\dim V$ generic hyperplanes.
%% %
%% \begin{example} 
%% An algebraic set of dimension zero is a finite set, with degree equal to its cardinality.
%% \end{example}
%% %
%% \begin{example}
%% An algebraic set of dimension 1 is a curve, with degree equal to the number of intersection points with a generic hyperplane.
%% \end{example}
%% %
%% %
%% %

%% Let $V \subset \C^n$ be a $d$-equidimensional algebraic set.  The
%% point $x \in V$ is a \textit{regular point} if $\dim (T_xV) = d.$
%% Otherwise $x$ is a \textit{singular point}. We let $\reg(V)$ and
%% $\sing(V)$ respectively denote the regular and singular points of $V$.

%% %
%% \subsubsection{Critical points of polynomial mappings}
%% %
%% Let $\phi : V \rightarrow \C^m$ be a polynomial mapping. A \textit{critical point} of $\phi$ is a regular point of $V$ with $d_x \phi(T_xV) \not = \C^m,$ where $d_x \phi$ is the differential of $\phi$ at $x.$ A \textit{critical} value of $\phi$ is the image of a critical point by $\phi$.
%% %
%% \subsection{The zariski-tangent space}
%% %
%% Denote by $\grad_{\xb}(f)$ the evaluation of the gradient vector of $f$ in $\C[X_1,\hdots,X_n]$ at $\xb$ in $\C$.
%% The \textit{Zariski-tangent space} to $V$ at $\xb \in V$ is the vector space $T_{\xb}V$ defined by the equations $\grad_{\xb}(f) \cdot \textit{\textbf{v}}=0$ for all polynomials $f$ that vanish on $V$. For a polynomial system $\textit{\textbf{f}}=(f_1,\hdots,f_N)$ in $\C[X_1,\hdots,X_n]$, denote by $\jac(\textit{\textbf{f}})$ the Jacobian matrix.
%% The following is a direct consequence of \cite[Corollary 16.20]{ECA}.
%% %
%% \begin{proposition}
%% If $V \subset \C^n$ is a $d$-equidimensional algebraic set with ideal $I(V)=\langle f_1,\hdots,f_N \rangle$, then at any point $\textbf{x}$ of $\reg(V),$ $\jac_\textbf{x}(\textbf{f})$ has full rank $n - \dim(V)$ and the kernel of $\jac_\textbf{x}(\textbf{f})$ is equal to $T_\textbf{x}V.$ 
%% \end{proposition}
%% %
%% \subsection{Polar varieties}
%% %
%% %
%% \subsection{Changes of variables}
%
Consider $A \in \C^{n^2}, f \in \C[X_1,\hdots,X_n], \xb \in \mathbb{C}^n,$ and $V \subset \C^n$. We denote by 
$V^A$ the set $\{x^A~|~\xb \in V\}, f^A$ the polynomial $f(A \xb)$ and $\xb^A$ denotes $A^{-1}\xb$. Notice $f^A(\xb^A) = f(A A^{-1}\xb)=f(\xb).$
%
%
%
%
%
%
%
%
\section{Genericity properties}
%
Consider $f \in \mathbb{Q}[X_1,\hdots,X_n]$ with total degree $D$.  
We say that $f$ satisfies $\textbf{H}$ if 
%
\begin{enumerate}
    \item $f$ is squarefree;
    \item $V(f)$ is smooth.
\end{enumerate}
%
An equidimensional algebraic set $X$ of dimension $d$ is in \textit{Noether position} for $\pi_d$ when the extension 
\[
\C[X_1,\hdots,X_d] \rightarrow \C[X_1,\hdots,X_n]/I(V)
\] 
is injective and integral.  Then, for any $\xb \in \C^d,$ the fiber 
$V \cap \pi_d^{-1}(\xb)$ has dimension zero. For $i\in\{1,\hdots,n\}$, and for a randomly chosen point $p_{i-1} = (x_1,\hdots,x_{i-1}) \in \Q^{i-1}$, we say that $f$ satisfies $\textbf{H}_i^{'}$ if 
%
\begin{enumerate}
\item $\I(\pi_i,V)$ is radical;
\item either $W(\pi_i,V)$ is empty or $(i-1)$-equidimensional;
\item $W(\pi_i,V)$ is smooth;
\item $\langle X_1-x_1,\hdots,X_{i-1}-x_{i-1}, f,  \frac{\pa f}{\pa X_{i+1}},\hdots,\frac{\pa f}{\pa X_n}\rangle $ is radical; 
\item $V(X_1-x_1,\hdots,X_{i-1}-x_{i-1}, f,  \frac{\pa f}{\pa X_{i+1}},\hdots,\frac{\pa f}{\pa X_n})$ is finite;
\item $V(X_1-x_1,\hdots,X_{i-1}-x_{i-1}, f,  \frac{\pa f}{\pa X_{i+1}},\hdots,\frac{\pa f}{\pa X_n})$ is smooth;
\item either $W(\pi_i,V)$ is empty or in Noether position for $\pi_{i-1}$.
\end{enumerate}
%
\begin{theorem}
Suppose that $f \in \Q[X_1,\hdots,X_n]$ satisfies \textbf{H}. Then there exists a Zariski closed set $\D \subset \mathbb{C}^{n^2}$ of degree at most $D^{\textrm{\td}}$ such that, if $\textbf{A}$ is chosen from  $\C^{n^2}-\Delta$ ( $\Q^{n^2}-\Delta$?) then $f^A$ satisfies $\textbf{H}_i^{'}$ for all $i \in \{1,\hdots,n\}.$
\end{theorem}
%
\begin{corollary} 
Fix $S \subset \mathbb{Q}$ with $|S| \geq \epsilon^{-1} D^{\textrm{\td}}$ and $\epsilon > 0$. Then for $A\in S^{n^2}$ chosen randomly, the probability that $f^A$ satisfies $H_i^{'}$ for all $i \in \{1,\hdots,n\}$ is at least $1-\epsilon.$
\end{corollary}
%
%
%
%
\section{Weak Transversality}
